Recovering from population extinction in the Animal Life Cycle Algorithm (ALCA)

### 2022-09-23 paper.tex
Updated Improvements
Work in progress.

The main contribution of this paper is to demonstrate that it is possible to
evolve a population of individuals, similar to a Genetic Algorithm (GA), using
a distributed, parallel, and asynchronous methodology by the algorithm
implementation and testing. We improve the algorithm behaviour by solving
population extinction caused by nature's pressure in the Animal Life Cycle
Algorithm with one of the most promising alternatives: the projection of the
Elite towards the Champion, where we compare the results obtained with the
latter and the random use of the more traditional alternatives (mutation and
crossover), using classic benchmark functions for optimization for comparison.

We present the description of the following sections in our paper. First, we
illustrate our algorithm model, the problem found with the occasional
population extinction and our proposed solution in
section~\ref{section.proposal}, followed by our experimental configuration and
results in section~\ref{section.experiments}, where we continue to analyze and
describe some of our research findings in section~\ref{section.discussion}. We
finalize by presenting some inferences based on our experiments results in
section~\ref{section.conclusions}.

@@

Our inspiration for this algorithm was found on the study and observation of
nature, using general abstraction to identify what most animal species have in
common: the life cycle. We present an algorithm inspired by the biological
life-cycle of animal species, which consists of several stages
\cite{read1968system}: birth, growth, reproduction, and death. The combination
of those processes results in evolution. As in nature, we intend to execute all
these stages in parallel and asynchronously on a constantly evolving
population.

@@

In this article, we test various strategies to generate a new population to
replace individuals that have already died (or were discarded) due to their
aptitude, as happens in the life cycle. We previously used to generate a new
random set of solutions, with the cost of losing all evolution knowledge. The
strategy we describe on this paper uses elites to take advantage of the
historical evolution, especially the best individuals, where we look for the
offspring of the elite set to be somewhat close to them. We use the golden
ratio to compute the distance between new individuals in the population, making
an improved distribution.

We propose one promising alternative to restart the population, to solve
extinction caused by nature's pressure in the Life Cycle algorithm: the
projection of the historically best-found candidate solutions (Elite) towards
the best-found candidate solution (Champion). This article proposes some
strategies to carry out this operation, where the presented distribution
strategy obtained best results.

\subsubsection{How does it work?} Based on the historically best candidate
solution and the elite set, we use the Golden Ratio (shown in
Figure~\ref{fig.golden_ratio}) to compute the distance between new individuals
in the population, making an improved distribution. It randomly selects a
solution from the elite set and uses the Golden Ratio to compute multiple
solutions against the best found candidate.

@@

This strategy is how we can calculate the projection of the historically
best-found candidate solutions (Elite) towards the best-found candidate
solution (Champion). To illustrate this concept we use
Figure~\ref{fig.elite_projection}, placing the Elite solution on the far right
and the Champion solution on the left corner, where a higher number of new
solutions will be closer to the Champion.

@@

This process generates new solutions that will be the new population set. It
uses the Golden Ratio to progressively aproximate from an Elite solution
towards the Champion. This strategy facilitates to find apparently hidden
solutions, having an exploitation emphasis.
Figure~\ref{fig.elite_projection_swarm} displays the projection of five Elite
solutions (on the corners) towards the Champion (on the center).

@@

ALTO / HASTA AQUI VAMOS.

>>>>

The main contribution of this paper is to demonstrate that it is possible to
evolve a population of individuals, similar to a Genetic Algorithm (GA), using
a distributed, parallel, and asynchronous methodology by algorithm
implementation and testing. We improve the algorithm behavior by solving
population extinction caused by nature's pressure in the Animal Life Cycle
Algorithm with one of the most promising alternatives: the projection of the
Elite towards the Champion, where we compare the results obtained with the
latter and the random use of the more traditional alternatives (mutation and
crossover), using classic benchmark functions for optimization for comparison.

We present the description of the following sections in our paper. First, we
illustrate our algorithm model, the encountered problem with the occasional
population extinction, and our proposed solution in
section~\ref{section.proposal}, followed by our experimental configuration and
results in section~\ref{section.experiments}, where we continue to analyze and
describe some of our research findings in section~\ref{section.discussion}. We
finalize by presenting some inferences based on the results of our experiments
in section~\ref{section.conclusions}.

@@

Our inspiration for this algorithm was found in the study and observation of
nature, using general abstraction to identify what most animal species have in
common: the life cycle. We present an algorithm inspired by the biological
life-cycle of animal species, which consists of several stages
\cite{read1968system}: birth, growth, reproduction, and death. The combination
of those processes results in evolution. As in nature, we intend to execute all
these stages in parallel and asynchronously on a constantly evolving
population.

@@

In this article, we test various strategies to generate a new population to
replace individuals that have already died (or were discarded) due to their
aptitude, as happens in the life cycle. We previously used to generate a new
random set of solutions, with the cost of losing all evolution knowledge. The
strategy we describe in this paper uses elites to take advantage of the
historical evolution, especially the best individuals, where we look for the
offspring of the elite set to be somewhat close to them. We use the golden
ratio to compute the distance between new individuals in the population, making
an improved distribution.

We propose one promising alternative to restart the population, to solve
extinction caused by nature's pressure in the Life Cycle algorithm: the
projection of the historically best-found candidate solutions (Elite) towards
the best-found candidate solution (Champion). This article proposes some
strategies to carry out this operation, where the presented distribution
strategy obtained the best results.

\subsubsection{How does it work?} Based on the historically best candidate
solution and the elite set, we use the Golden Ratio (shown in
Figure~\ref{fig.golden_ratio}) to compute the distance between new individuals
in the population, making an improved distribution. It randomly selects a
solution from the elite set and uses the Golden Ratio to compute multiple
solutions against the best-found candidate.

@@

This strategy is how we can calculate the projection of the historically
best-found candidate solutions (Elite) towards the best-found candidate
solution (Champion). To illustrate this concept we include
Figure~\ref{fig.elite_projection}, placing the Elite solution on the far right
and the Champion solution on the left corner, where a higher number of new
solutions will be closer to the Champion.

@@

This process generates new solutions that will be the new population set. It
uses the Golden Ratio to progressively approximate from an Elite solution to
the Champion. This strategy facilitates to find apparently hidden solutions,
having an exploitation emphasis. Figure~\ref{fig.elite_projection_swarm}
displays the projection of five Elite solutions (on the corners) towards the
Champion (in the center).

@@

As a proof-of-concept, we implemented the algorithm with Docker containers,
where we compare the results obtained with the projection of the Elite towards
the Champion strategy and the random use of the more traditional alternatives
(such as mutation and crossover), using classic benchmark functions for
optimization as implemented by the DEAP (Distributed Evolutionary Algorithms in
Python) library \cite{fortin2012deap}.

The alternatives to restart this new population are the creation of a new set
of candidate solutions based on either: The mutation with uniform modification
from the Elite or the historically best-found solution (Champion); Crossing the
Champion with the Elite; The projection of the Elite towards the Champion;
Based on random use of the previous alternatives.

\subsection{ALTO / HASTA AQUI VAMOS}
